---
title: "Commerce Comission Lines Company Data"
subtitle: "Looking for load control data"
author: "Ben Anderson, b.anderson@soton.ac.uk"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
always_allow_html: yes
bibliography: '`r path.expand("~/bibliography.bib")`'
---


```{r knitrSetup, include=FALSE}
rmdParams <- list()

rmdParams$startTime <- proc.time() # used at the end

knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo

# libraries
library(data.table)
library(flextable) 
library(ggplot2)
library(plotly)
library(readxl)

# > functions ----
makeFlexTable <- function(df, cap = "caption"){
  # makes a pretty flextable - see https://cran.r-project.org/web/packages/flextable/index.html
  ft <- flextable::flextable(df)
  ft <- colformat_double(ft, digits = 1)
  ft <- fontsize(ft, size = 9)
  ft <- fontsize(ft, size = 10, part = "header")
  ft <- set_caption(ft, caption = cap)
  return(flextable::autofit(ft))
}

# paths

rmdParams$dPath <- path.expand("~/Dropbox/Work/Otago/data/comCom/") # change for you

```

# Introduction

Looking for load control data

# Inquiry into 'new' technologies

Data source: https://comcom.govt.nz/regulated-industries/electricity-lines/electricity-distributor-performance-and-data/impact-of-emerging-technologies-in-monopoly-parts-of-electricity-sector

```{r loadNewTechData}
df <- readxl::read_xlsx(paste0(rmdParams$dPath, "Electricity-distribution-businesses-emerging-technology-data-10-October-2018.xlsx"),
                        sheet = "Database")

dt <- data.table::as.data.table(df)

names(dt)
makeFlexTable(head(dt))

makeFlexTable(dt[Category %like% "control", .(nRows = .N), keyby = .(Category, Description)])
```

Check who may not have returned specific data items under ripple/other load control...

```{r extractLoadControl}
rc_dt <- dt[Description %like% "Estimated number of ICPs" |
              Description %like% "Number of ICPs" |
              Description %like% "Load control capacity from ripple control" |
              Description %like% "Load control capacity from other load control"]

message("Any NA data?")

rc_dt[, numValue := as.numeric(Value)]

naT <- rc_dt[is.na(numValue)]

ft <- makeFlexTable(naT[,.(EDB, Year, Description, Value)])
set_caption(ft, "Rows with missing data against ripple or other load control entries")
```


> So some EDBs returned unknown to both in 2014 (Electricity Invercargill, OtagoNet, The Power Company), Vector returned unknown to ripple control capacity in all years but must have reported numbers.

## Numbers of ICPs

Cross-check against \@ref(tab:extractLoadControl) for potentially missing EDBs.

Figure \@ref(fig:plotNICP) shows declared number of ICPs with load control over time.

```{r plotNICP, fig.cap="Declared number ICPs with load control over time"}
t <- rc_dt[Description == "Estimated number of ICPs with ripple control" |
        Description == "Number of ICPs with other load control", .(nICPs = sum(as.numeric(Value), na.rm = TRUE)),
      keyby = .(Description, Year)]

makeFlexTable(t)

plotDT <- rc_dt[Description == "Estimated number of ICPs with ripple control" |
        Description == "Number of ICPs with other load control", .(nICPs = sum(as.numeric(Value), na.rm = TRUE)),
      keyby = .(Description, Year, EDB)]

p <- ggplot2::ggplot(plotDT, aes(x = Year, y = nICPs, color = EDB, group = EDB)) +
  geom_line() +
  facet_grid(Description ~ .)
  
p
```


Attempt to plotly the plot for interaction - try hovering:

```{r plotlyVersionNICPs}
plotly::ggplotly(p)
```

But check  values - are they 'clumped'? Or do they look realistic? Some of those look suspiciously flat...

```{r check_cap}
ggplot2::ggplot(plotDT, aes(x = nICPs)) +
  geom_histogram() +
  facet_grid(Year ~ Description, scales = "free")
```

Numbers do look a bit '[heaped](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5976260/)'...

## Capacity

Cross-check against \@ref(tab:extractLoadControl) for potentially missing EDBs.

> i.e. Vector in all years

Figure \@ref(fig:extractLoadControlSumMW) shows load control capacity over time (MW).

```{r extractLoadControlSumMW, fig.cap="Declared load control capacity over time (MW)"}
rc_dt <- dt[Description %like% "ripple control" |
              Description %like% "load control"]

t <- rc_dt[Description == "Load control capacity from ripple control" |
        Description == "Load control capacity from other load control", .(sumCapacity = sum(as.numeric(Value), na.rm = TRUE)),
      keyby = .(Description, Year)]

makeFlexTable(t)

plotDT <- rc_dt[Description == "Load control capacity from ripple control" |
        Description == "Load control capacity from other load control", .(sumCapacity = sum(as.numeric(Value), na.rm = TRUE),
                                                                          nValues = .N),
      keyby = .(Description, Year, EDB)]
# check nValues = 1 (if so we don't actually need to sum)

p <- ggplot2::ggplot(plotDT, aes(x = Year, y = sumCapacity, color = EDB, group = EDB)) +
  geom_line() +
  facet_grid(. ~ Description, scales = "free") +
  labs(y = "Capacity (MW)") +
  theme(legend.position = "bottom")

p
```

Attempt to plotly the plot for interaction - try hovering:

```{r plotlyVersionMW}
plotly::ggplotly(p)
```

But check capacity values - are they 'clumped'? Or do they look realistic? Some of those look suspiciously flat...

```{r check_nICP}
ggplot2::ggplot(plotDT, aes(x = sumCapacity)) +
  geom_histogram(binwidth = 2) +
  facet_grid(Year ~ Description, scales = "free")
```
Does look a bit like there is clumping/heaping - so some of the values for ripple control may be 'estimated & rounded' rather than calculated from base data.

# Lines companies performance data

Data Source: https://comcom.govt.nz/regulated-industries/electricity-lines/electricity-distributor-performance-and-data/information-disclosed-by-electricity-distributors

Loading takes time - large .xlsm file...

> skip this section for now

```{r loadPerfData, warning=FALSE, eval = FALSE}
df <- readxl::read_xlsx(paste0(rmdParams$dPath, "/Electricity-distributers-information-disclosure-data-2013-2020.xlsm"),
                        sheet = "Database")

dt <- data.table::as.data.table(df)
names(dt)

makeFlexTable(head(dt))
```

Search for 'load' / 'Load' - maybe this didn't work too well.

```{r searchForLoad, eval = FALSE}
t_dt <- dt[Category %like% "load" | 
             `Sub-category` %like% "load" |
             Description %like% "load" |
             Category %like% "Load" | 
             `Sub-category` %like% "Load" |
             Description %like% "Load"]

makeFlexTable(t_dt[, .(count = .N), keyby = .(Category, `Sub-category`, Description)])
```

Anything there?

Search for 'control'

```{r searchForControl, eval = FALSE}
t_dt <- dt[Category %like% "control" | 
             `Sub-category` %like% "control" |
             Description %like% "control" ]

makeFlexTable(t_dt[, .(count = .N), keyby = .(Category, `Sub-category`, Description)])
```

Anything there?

# R environment

```{r check runtime, include=FALSE}
# within Rmd timing
t <- proc.time() - rmdParams$startTime
elapsed <- t[[3]]
```

Report generated in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * data.table [@data.table]
 * flextable [@flextable]
 * ggplot2 [@ggplot2]
 * knitr [@knitr]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

```{r runAll}
```

# References

